Predicting if barell are lifted correctly
=========================================

This is my submission to the Pratical Machine Learning Course Project. I kept it very simple, so it should be fast to read :).

Required libraries
==================

```{r}
library(caret)
```

Data loading and processing
===========================

First let's load the data (the files are included in the repository):

```{r}
# columns with incorrect values are set to nan
all_training <- read.csv('pml-training.csv', na.strings = c("", "NA", "#DIV/0!") )
all_testing <- read.csv('pml-testing.csv', na.strings = c("", "NA", "#DIV/0!") )
```

Next, we have to remove the first 7 columns as they are of little use.
```{r}
all_training = all_training[,-(1:7)]
all_testing = all_testing[,-(1:7)]
```

Removing all columns where there are at least one NA value (it might be a little extreme, but it is simpler like that).
```{r}
removeNaColumns     <- function(x) { x[, colSums(is.na(x)) == 0 ] }
all_training <- removeNaColumns(all_training)
all_testing <- removeNaColumns(all_testing)
```


Let's then separate our training data into two parts, for cross validation.
```{r}
partition  <- createDataPartition(all_training$classe, p=.6, list=FALSE)
raw_training <- all_training[partition,]
raw_validation  <- all_training[-partition,]
```

And we then apply PCA to reduce the number of dimensions:
```{r}
preProc = preProcess(raw_training[,-length(raw_training)], method="pca", thresh=0.9)
training <- predict(preProc, raw_training[,-length(raw_training)])
validation <- predict(preProc, raw_validation[,-length(raw_validation)])
```

Training
========
We trained the model on Random forest:
```{r}
modFit <- train(raw_training$classe~ .,data=training,method="rf",prox=FALSE)
```

Results
=======
We can now predict our validation set:
```{r}
predicted = predict(modFit, validation)
```

The overall success rate is (that should be near our out of sample error):
```{r}
sum(predicted == raw_validation$classe) / length(predicted)
```

Here is the confusion matrix:
```{r}
confusionMatrix(predicted, raw_validation$classe)
```
